# Sitzung 06
## Speichern und laden der Datenstrukturen in R

    save(some.data.name, file = sprintf("some_dir/%s-%s.RData", dates[1], dates[2]))
    load(sprintf("./precomputed/%s-%s.RData", dates[1], dates[2]))


## Frames speichern:
  
    write.table(x, file=, sep='\t') # Siehe `?write.table`

## Frames einlesen:

    # Mit Standardwerten spielen, `?read.table`
    # x <- read.table(file='', header=(TRUE|FALSE), row.names = 1)

## `fix()`: Eingebauter Editor

    fix(x)

## Deskriptiv statistics
```{r}
library(moments)

random.vector <- rnorm(100)
str(random.vector)
summary(random.vector)
mean(random.vector)
median(random.vector)
quantile(random.vector, c(0.25, 0.5, 0.75))
sd(random.vector)
var(random.vector)
mad(random.vector)
skewness(random.vector)
kurtosis(random.vector)
    
```

## Graphische Darstellung
```{r Temperaturvergleich, highlight = TRUE, results = 'hide', echo = TRUE}
# some vectors
monate <- c('Jan', 'Feb', 'Mar', 'Apr', 'Mai', 'Juni', 'Juli', 'Aug', 'Sep', 'Okt', 'Nov', 'Dez')
stadt1 <- c(-5, -12, 5, 12, 15, 18, 22, 23, 20, 26, 8, 1)
stadt2 <- c(6, 7, 8, 9, 10, 12, 16, 15, 11, 9, 8, 7)



#pdf('out.pdf')
plot(c(1:12), stadt1, type='b', col='dark red',
     xlim=c(0,12), ylim=c(min(stadt1, stadt2),
                          max(stadt1, stadt2)),
     xlab='Monate', ylab='Temperatur im Schnitt',
     main='Temperaturvergleich',
     axes = FALSE)

axis(1, at=1:12, labels=monate, las=2)
axis(2, at=sort(c(stadt1, stadt2)))


#mtext('Monate', side=1)
#mtext('Temperatur im Schnitt', side=2)
points(c(1:12), stadt2, type='b', col='green')
#grid(nx=13)
grid(NA, 39)
abline(h=0)
text(11, 25, 'Stadt 1')
text(7, 10, 'Stadt 2')

#box()
#dev.off()
```
```{r Kuchendiagramme}
library(plotrix)
POS <- c(3,4,5,3,1,2,2)
labs <- c('A', 'M', 'N', 'P', 'Part', 'V', 'Adj')
names(POS) <- c('A', 'M', 'N', 'P', 'Part', 'V', 'Adj')

par(mfrow=c(1,2))

pie(POS, main="Kuchendiagramm")

pie3D(POS,
      main="POS-Verteilung",
      labels=names(POS),
      explode=0.1
      )
```
```{r combined_graphics}
freq <- c(rep(0, 1), rep(1000, 2), rep(2000, 3), rep(3000, 4), rep(4000, 5), rep(5000, 5), rep(6000, 5), rep(7000, 15), rep(10000, 26), rep(15000, 26), rep(25000, 8), rep(50000, 1))

bins <- c(0, 1000, 2000, 3000, 4000, 5000, 6000, 7000, 10000, 15000, 25000, 50000)

par(mfrow=c(2,2))

par(fg='blue')
hist(freq, freq=FALSE, breaks=bins, right = FALSE, axes = FALSE, main = NULL, xlab = NULL, ylab = NULL)

axis(1,
     at=bins,
     labels=TRUE, # default
     tick=TRUE, # default
     las=2, # vertical rotation
     ps=1
     )

barplot.matrix <- read.csv('barplot.csv', header=TRUE) 

barplot(as.matrix(t(barplot.matrix)[2:3,]), col=heat.colors(2), beside=TRUE,
space=c(1,3), ylim=c(0,15), names.arg=barplot.matrix$Rang)
grid()
abline(0,0)

library(plotrix)
slices <- c(10, 12, 4, 16, 8)
names(slices) <- c('A', 'M', 'N', 'P', 'V')
pie(slices, main="Kuchendiagramm")
pie3D(slices, explode=0.1, labels=names(slices))

```

